Demography and linked selection in wild and domesticated maize

### 2015-09-01 manuscript.tex
edits

%Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}. However, we implemented a scan based on the H12 statistic, which is designed to identify both hard and soft sweeps with reasonable power \cite{garud2015}. Although the goal of this study was not to identify specific sweeps, it may be that the outlier sites distinguished by H12 are primarily composed of soft sweeps instead of hard. Similarly, previous studies scanning for evidence of positive selection in maize \cite{hufford2012} may have picked up primarily evidence of soft sweeps.

%We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution. In fact, the maize locus \emph{tga1} that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}. Surrounding \emph{tga1},  our data demonstrate a pattern consistent with a hard sweep. But, our data demonstrate that instances such as this are infrequent. This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and non-synonymous substitutions are clear suggesting hard-sweeps are abundant. However, it agrees with what has been seen for humans \cite{hernandez2011}.

@@

Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of background selection  \cite{enard2014}. 
Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability. 
Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}. 
Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation. 
Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}.

@@

\subsection{Demography of domestication}
Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.  
We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sties.  
By utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.

@@

%%%Shorten if we include MSMC
%Finally, MSMC \cite{schiffels2014} was employed to complement our model-based demographic inference. 
%Using six each of maize and teosinte
%The analyses were run for the following two groups: six maize haplotypes (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033) and six teosinte haplotypes (TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14).
%SNPs were called in ANGSD \cite{korneliussen2014} with the SNP p value set as $1e-6$.
%As the samples are highly inbred, each individual was treated as one haplotype.
%Only Sites with mapping quality bigger than $30$, base quality over $20$ and depth between half and twice of its mean depth were chosen for the analyses, and the residual heterozygous sites were removed from the analyses.
%In addition, as the maize genome contains a large body of repetitive regions \cite{schnable2009}, we used Heng Li's software SNPable (\href{<url>}{<http://lh3lh3.users.sourceforge.net/snpable.shtml>}) to mask out genomic regions, on which short sequencing reads are not uniquely mapped.
%The mappability mask file was generated by cutting the maize genome with $1bp$ shift each time into the $100bp$ single-end reads, which were then mapped back to the maize B73 reference genome \cite{schnable2009}.
%The sites with the majority of overlapping $100$-mers mapped uniquely without one mismatch were called the "SNPable" sites and used for the MSMC analyses.
%The pattern parameter $20\times2+20\times4+10\times2$ was used for the population size inference.

>>>>

@@

Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of background selection  \cite{enard2014}. 
Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability. 
Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}. 
Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation. 
Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}. \jri{no mention of our H12 results. I think that's OK?}

@@

\subsection{Demography of domestication}
Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.  
We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.  
By utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.

@@

%%%Shorten text here if we include MSMC
%Finally, MSMC \cite{schiffels2014} was employed to complement our model-based demographic inference. 
%Using six each of maize and teosinte
%The analyses were run for the following two groups: six maize haplotypes (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033) and six teosinte haplotypes (TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14).
%SNPs were called in ANGSD \cite{korneliussen2014} with the SNP p value set as $1e-6$.
%As the samples are highly inbred, each individual was treated as one haplotype.
%Only Sites with mapping quality bigger than $30$, base quality over $20$ and depth between half and twice of its mean depth were chosen for the analyses, and the residual heterozygous sites were removed from the analyses.
%In addition, as the maize genome contains a large body of repetitive regions \cite{schnable2009}, we used Heng Li's software SNPable (\href{<url>}{<http://lh3lh3.users.sourceforge.net/snpable.shtml>}) to mask out genomic regions, on which short sequencing reads are not uniquely mapped.
%The mappability mask file was generated by cutting the maize genome with $1bp$ shift each time into the $100bp$ single-end reads, which were then mapped back to the maize B73 reference genome \cite{schnable2009}.
%The sites with the majority of overlapping $100$-mers mapped uniquely without one mismatch were called the "SNPable" sites and used for the MSMC analyses.
%The pattern parameter $20\times2+20\times4+10\times2$ was used for the population size inference.
